---
title: "DATA5002 Project"
author: "HONGQING YU"
date: "2024-11-12"
output: html_document
---
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(plotly)
library(DT)
library(reshape2)
library(crosstalk)
library(dplyr)
```

## Main factors of diabetes and how to reduce the possibility of diabetes by controlling BMI
### Heatmap of all variables in the Diabetes Dataset

```{r echo=FALSE, message=FALSE, warning=FALSE}
diabetes_data <- read.csv("diabetes_dataset.csv", header = TRUE)
diabetes_data <- diabetes_data[diabetes_data$BMI != 0, ]

heatmap_data <- cor(diabetes_data)
ggplot(melt(heatmap_data), aes(Var1, Var2, fill = value)) + 
    geom_tile() +
    theme_minimal() 
```

### Boxplot of Glucose by Outcome
```{r echo=FALSE, message=FALSE, warning=FALSE}
diabetes_data$Outcome <- as.factor(diabetes_data$Outcome)

plot_ly(data = diabetes_data, 
               x = ~Outcome,
               y = ~Glucose,
               type = 'box',
               color = ~Outcome,
               colors = c("#6A9ACE","#715EA9"),
               boxmean = TRUE,
               boxpoints = 'all',  
               jitter = 0.3,       
               line = list(color = 'royalblue2'))


```


### Density Plot of Family History by Outcome
```{r echo=FALSE, message=FALSE, warning=FALSE}
#diabetes&faminly history

ggplot(diabetes_data, aes(x = DiabetesPedigreeFunction, fill = Outcome)) +
  geom_density(alpha = 0.5) +
  labs(x = "Family History", y = "Density", fill = "Outcome") +
  theme_bw() +
  theme(panel.border = element_blank())

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# fig <- plot_ly(data = diabetes_data, 
#                x = ~DiabetesPedigreeFunction, 
#                color = ~Outcome, 
#                colors = c("salmon", "lightseagreen"),
#                type = 'histogram', 
#                histnorm = 'density', 
#                opacity = 0.8) %>%
#   layout(title = 'Density Plot of Family History by Outcome',
#          xaxis = list(title = 'Family History'),
#          yaxis = list(title = 'Density'),
#          barmode = 'overlay')
# 
# fig

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
diabetes_data$age_group <- cut(
  diabetes_data$Age,
  breaks = c(seq(21, 71, by = 10), Inf),
  right = FALSE,
  labels = c("21-31", "32-41", "42-51", "52-61", "62-71", "72-81")
)

diabetes_data$BMI_group <- cut(
  diabetes_data$BMI,
  breaks = c(-Inf, 18.5, 25, 30, 40, Inf),  
  labels = c("Under Weight", "Normal", "Overweight", "Obese", "Extremely Obese"),  
  right = FALSE  
)
diabetes_data$preg_group <- cut(
  diabetes_data$Pregnancies,
  breaks = c(0, 4, 8, 12, Inf),  
  labels = c("None-3 children", "4-7 children", "8-11 children", "More than 11 children"),  
  right = FALSE  
)

```

### Outcome Count by Age, BMI Level and Number of Children
```{r echo=FALSE, message=FALSE, warning=FALSE}
diabetes_data2 <- diabetes_data %>%
  group_by(Outcome, age_group, BMI_group, preg_group) %>%
  summarise(count = n(), .groups = "drop")

diabetes_data2 <- diabetes_data2 %>%
  mutate(if_diabetes = ifelse(Outcome == 1, "diabetes", "healthy"))

d <- SharedData$new(diabetes_data2)
bscols(
  widths = c(2, NA),
  list(
    filter_checkbox("age_group", "Age Group", d, ~age_group, inline = FALSE),
    filter_select("BMI_group", "BMI Group", d, ~BMI_group, multiple = FALSE)
  ),
  list(
    plot_ly(d, x = ~Outcome, y = ~count, type = "bar", color = ~preg_group)%>%
      layout(
        #title = "Outcome Count by Age, BMI and Number of Children",
        xaxis = list(type = "category", title = "Outcome"),
        yaxis = list(title = "Count")
      )
  )
)



```




##                                           
##       




### BMI Level Distribution and Gender
```{r echo=FALSE, message=FALSE, warning=FALSE}
obesity_data <- read.csv("obesity_level.csv", header = TRUE)
obesity_data <- obesity_data %>%
  rename(Level = X0be1dad)

#level&gender
obesity_data2 <- obesity_data %>%
  group_by(Gender, Level) %>%
  summarise(count = n(), .groups = "drop")
d1 <- SharedData$new(obesity_data2)

bscols(
  widths = c(2, NA), 
  list(
    filter_checkbox("gender", "Gender", d1, ~Gender, inline = FALSE)
  ),
  list(
    plot_ly(d1, 
            labels = ~Level, 
            values = ~count,      
            type = 'pie',
            colors = "Set1") %>%
    layout(
           #title = "Obesity Level Distribution",
           showlegend = TRUE)
  )
)
```


### Obesity Level Distribution and Family History
```{r echo=FALSE, message=FALSE, warning=FALSE}

#level&Consumption of alcohol
obesity_data3 <- obesity_data %>%
  group_by(Level, CALC, family_history_with_overweight) %>%
  summarise(count = n(), .groups = "drop")
d2 <- SharedData$new(obesity_data3)

bscols(
  widths = c(2, NA), 
  list(
    filter_checkbox("family_history_with_overweight", "family history", d2, ~family_history_with_overweight, inline = FALSE),
    filter_select("CALC", "Consumption of alcohol", d2, ~CALC, multiple = FALSE)
  ),
  list(
    plot_ly(d2, 
            labels = ~Level, 
            values = ~count,      
            type = 'pie',
            colors = "Set1") %>%
    layout(
           #title = "Obesity Level Distribution",
           showlegend = TRUE)
  )
)


```














### Outcome Count by BMI Level and Trip Mode
```{r echo=FALSE, message=FALSE, warning=FALSE}
#level&trip mode
obesity_data3 <- obesity_data %>%
  group_by(Level, MTRANS) %>%
  summarise(count = n(), .groups = "drop")

plot_ly(obesity_data3, x = ~Level, y = ~count, color = ~MTRANS, type = "bar")

```

### Scatter Plot for Height and Weight
    This scatter plot shows the interval where each weight category is located distictly. You can bring in your own height and weight to determine your range.
```{r echo=FALSE, message=FALSE, warning=FALSE}

fig <- plot_ly(data = obesity_data, 
               x = ~Height, 
               y = ~Weight, 
               color = ~Level, 
               colors = "Set1", 
               opacity = 0.8, 
               type = 'scatter',        
               mode = 'markers')        

fig

```



































